---
title: "CNV Frequency Analysis in Cancer"
author: "Marieluisa Buchegger"
date: "2025-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Install packages and Load libraries

```{r eval =FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.20")

BiocManager::install(c("pgxRpi", "GenomicRanges"))

```

## Load Packages
```{r }
library(pgxRpi)
library(GenomicRanges)
```

## CNV Frequency Analysis in GLlioblastoma

I'd like to study Glioblastoma .In Progenetix, the corresponding NCIt code is "NCIT:C3058".
Glioblastoma is one of most frequent malignant brain tumors. It is a highly agrressive cancer that can lead to death withinin a short time if left untreated. Even with treatment, glioblastoma often recurs and is associated with a poor prognosis. Research indicates that genetic alterations play a key role in the development of glioblastoma.
Source: link[(https://my.clevelandclinic.org/health/diseases/17032-glioblastoma)]

## Access CNV frequency data from samples with glioblastoma
```{r}
freq <- pgxLoader(type ="cnv_frequency", filters = "NCIT:C3058")

```

To access the CNV frequency data from this object:
```{r}
freq[["NCIT:C3058"]]
```

To get the metadata, use 'mcols' function:
```{r}
mcols(freq)
```

Visualize the frequency data - By entire genome
```{r, fig.width= 12, fig.height = 6}
pgxFreqplot(freq)
```
Visualize the frequency data - By Chromosome
```{r, fig.width= 12, fig.height = 6}
pgxFreqplot(freq, chrom = c(7, 9, 10, 13, 20), layout = c(3, 2))
```


## Frquency Analysis
According to the plot, using a threshold of 25% for duplication and deletion. 7p, 7q, 19p, and 20q are frequently duplicated, 9p, 10p, 10q, and 13q are frequently deleted.

A previous study by Crespo et al. (2010) ([link (https://doi.org/10.1186/1753-6561-4-S2-P53)]) analyzed copy number gains and losses in a group of 35 glioblastoma cases utilizing SNP arrays. They reported gains in chromosome 7 in 97% of cases, deletions in chromosome 9 in 60%, and deletions in chromosome 10 in 83%. While my CNV frequency analysis reveals a similar pattern across these chromosomes, the specific percentages differ quite a bit (Chromosome 7: ~50% gains, Chromosome 9: ~30% losses, Chromosome 10: ~50% losses)

This may be attributed to the broader data sources used by the Progenetix database, which includes data from SNP arrays, array CGH, as well as whole-genome and whole-exome sequencing studies. This paper also only includes 35 cases, whereas our data has over 4000 samples. Including more diverse platforms and more samples could lead to a more heterogeneous dataset, potentially lowering the overall frequencies for certain events.

Additionally, in my analysis, chromosomes 19p and 20q showed gains, and 13q showed losses—these just surpassed my chosen threshold. It is possible that Crespo et al. applied a higher cutoff for reporting CNVs, which may explain why these alterations were not mentioned in their results.


