---
title: "bio392_cnv_freq"
author: "Serena Galli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#### Melanoma

**Melanoma** (*NCIT code = C3224*) is a type of cancer that develops from the pigment-producing cells known as melanocytes. Melanomas typically occur in the skin, but may rarely occur in the mouth, intestines, or eye (so-called uveal melanoma). Risk factors for melanoma include excessive sun exposure, a history of sunburns, fair skin, a family history of melanoma, and having many moles or atypical moles. 


#### Load the required libraries and access the data

```{r}
library(pgxRpi)
freq <- pgxLoader(type='frequency', output='pgxfreq', filters= "NCIT:C3224")
```

The metadata information is 

```{r}
freq$meta
```

The actual data looks like this: 

```{r}
freq$data$`NCIT:C3224`[c(1:3),] #show first three rows
```


#### Visualize the CNV frequency pattern

```{r}
pgxFreqplot(freq) #yellow = gains, blue = losses
```

```{r}
pgxFreqplot(freq, chrom=c(4,8, 17, 22)) #visualize 4 chromosomes in detail
```


#### Analysis: Which chromosomes display frequent gains or losses?
Using a cut-off of 25%, the following gains and losses are detected:  
gains: 1q, 6p, 7p, 7q, 8q, 17q (partially), 20p, 20q and 22p.  
losses: 4q (borderline), 6q, 9p, 9q, 10p, 10q and 11q. 


#### Comparison with a published paper

In a [paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1892437/) from 2003, a cohort of 186 melanocytic tumors encompassing both melanomas (n=132) and benign nevi (n=54) was studied. Amongst other things, they compared the presence of chromosomal aberrations, as assessed by CGH, between benign and malignant lesions and between different melanoma subtypes.  
Again relying on a cut-off of 25%, following gains and losses can be observed in Figure 1A (acral and non-acral melanoma):  
gains: 1q, 5p (acral melanoma), 6p, 7p, 7q, 8q (borderline), 17q , 20q (acral) and 22p.  
losses: 6q, 9p, 9q, 10p, 10q and 11q (acral, borderline).  

Following differences between the two ressources can be noted:  
gains: 5p (only in the paper), 20p (only in the progenetix sample collection, present at a lower percentage in the paper), 22q (only in the progenetix sample collection) 
losses: 4q (only in the progenetix sample collection, present at a lower percentage in the paper). 

The gains and losses observed in the two datasets overlap quite well. In my view, the most notable difference regards the gain of 5p, which seems to be a specific feature of acral melanoma, which was analysed separately in the paper (22/132 were acral melanomas). In the progenetix melanoma collection, no 'acral melanoma' category exists but given the fact that acral melanomas account for around 1-3% of all melanomas, one can guess that this subtype is underrepresented in the progenetix resource and therefore the frequency of this subtype-specific gain lies below 25%. 




