---
title: "RMarkdown toy example"
author: "Hangjia Zhao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Documentation

This is an **example** of R Markdown document. More details see the [documentation](http://rmarkdown.rstudio.com).

Simple dot points:

* Point 1
* Point 2
* Point 3

and numeric dot points:

1. Number 1
2. Number 2
3. Number 3

and nested dot points:

* A
    * A.1
    * A.2
* B
    * B.1
    * B.2

To denote a word as code or special parameters, use \` \`, such as `plot`.

## Code and results

### Load library

```{r}
library(datasets)
```

### Show the data

```{r}
head(iris)
```

### Plots

#### simple

```{r}
x <- c(1,2,3)
y <- c(2,4,6)
plot(x,y)
```

#### change figure size

```{r , fig.width=3, fig.height=3}
plot(x,y)
```

## R Code chunk features

When `echo` is set to false

```{r echo=FALSE}
5+5
```

When `eval` is set to false

```{r eval=FALSE}
5+5
```

Name a chunk

```{r, echo=FALSE}
Sys.sleep(10)
```

```{r special-chunk, echo=FALSE}
Sys.sleep(10)
```


