---
title: "CNV Frequencies - Invasive Breast Carcinoma"
author: "Evin Capan"
output: html_document
date: "2023-09-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tumor information

**Invasive Breast Carcinoma:** [The American Cancer Society](https://www.cancer.org/cancer/types/breast-cancer/about/types-of-breast-cancer/invasive-breast-cancer.html#:~:text=Breast%20cancers%20that%20have%20spread,carcinoma%20and%20invasive%20lobular%20carcinoma.) defines Invasive Breast Carcinoma as breast cancer that has spread into surrounding breast tissue. Most breast cancers are invasive, but there are different types of Invasive Breast Cancer. The two most common are Invasive Ductal Carcinoma (=IDC, 8 in 10 invasive breast carcinomas) and Invasive Lobular Carcinoma (ILC, 1 in 10 invasive breast carcinomas). Women are at a higher risk of developing invasive breast carcinoma than men. Potentially contributing to this risk are: Age, Genetic Factors (BRCA1/2), Hormonal influences, Lifestyle Factors and more.

The NCIt code is C9245.


## Load library

```{r}
library(pgxRpi)
```



## Access data

```{r}
freq <- pgxLoader(type='frequency',output='pgxfreq',filters='NCIT:C9245')
```


The metadata information is

```{r}
freq$meta
```


The data looks like this

```{r}
freq$data$`NCIT:C9245`[c(1:3),]
```



# Visualization

```{r}
pgxFreqplot(freq)
```



```{r}
pgxFreqplot(freq,chrom=c(1,8,16,17))
```



# Analysis

According to the plot, there are frequent duplications on chromosome 1q, 8q, 16p, 20q and frequent deletion on chromosomes 4p, 4q, 6q, 8p, 9p, 9q, 13q, 16q, 17p, 18p, 18q, 19p, 22q, Xp, Xq (25% as a cut-off). One [study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7780029/) detected that the most frequently altered regions were duplications on 1q, 3q and 8q and deletions on 16q and 17p.
When comparing both results, overlaps can be seen on: 1q, 8q for duplications; 16q, 17p for deletions.

The TP53 gene, which encodes the tumor suppressor protein p53, is on the short arm of chromosome 17 (17p). Deletions on 17p could increase the risk of Invasive Breast Carcinoma.