```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Install package

```{r}
if (!require(devtools)){
    install.packages('devtools')
}
devtools::install_github('progenetix/pgxRpi')
library(pgxRpi)
```

## Step2: Search esophageal adenocarcinoma NCIt code

```{r}
# query: esophageal adenocarcinoma, type of data: NCIt code, portal: progenetix.org - https://progenetix.org/service-collection/ontologymaps/

# => NCIT:C4025
```

## Step3: Access the CNV frequency data from samples with esophageal adenocarcinoma

```{r}
freq <- pgxLoader(type='frequency', output='pgxseg',filters='NCIT:C4025',
                  codematches=T)
```

### The retreived data is an object contaning two slots `meta` and `data`.

The `meta` slot looks like this (contains metadata, like sample count):

```{r}
freq$meta
```

The `data` slot includes two matrices.

```{r}
names(freq$data)
```

In this case, these two matrices are the same. The columns gain_frequency and loss_frequency are of primary interest.

```{r}
head(freq$data$`NCIT:C4025`)
```

```{r}
dim(freq$data$`NCIT:C4025`)
```

## Step4: Visualize data

### By genome

```{r}
pgxFreqplot(freq)
```

### By chromosome

```{r}
pgxFreqplot(freq,chrom = 20)
```

## Step5: Analyse the data

According the plot, we can see frequency gains on chromosome 7p, 8q, 20p, 20q 
and frequency losses on chromosome 4p, 4q, 5q, 9p, 17p, 18q, 21q.

There is a literature where the findings are consistent with the majority of mine. Here is the [paper-link](https://www.nature.com/articles/modpathol2008150).

A more detailed use case see this [link](https://htmlpreview.github.io/?https://github.com/progenetix/pgxRpi/blob/main/vignettes/Introduction_2_loadfrequency.html).
