---
title: "cnv-freq-analysis"
author: "Megumi Mizoguchi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages
Bioconductor
```{r eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
library(BiocManager)
BiocManager::install(version = "3.20")

BiocManager::install(c("pgxRpi", "GenomicRanges"))
```

## Load packages

```{r, message=FALSE}
library(pgxRpi)
library(GenomicRanges)
```

## Disease description

I'd like to study **Lung Squamous Cell Carcinoma**. In Progenetix, the corresponding NCIt code is **"NCIT:C3493"**.

## Access CNV frequency data from samples with lung squamous cell carcinoma

```{r}
#freq <- pgxLoader(type = "cnv_frequency", filters = "NCIT:C3493")
#Due to an issue of R in my PC, biosample information could not be retrieved.
#Error in curl::curl_fetch_memory(url, handle = handle) : 
#  Failure when receiving data from the peer [progenetix.org]: Recv failure: Connection reset by peer


#Finally, getting data via Python → load into R worked

#reticulate::py_install("requests", pip = TRUE)
#library(reticulate)
#requests <- import("requests")
#res <- requests$get("https://progenetix.org/services/intervalFrequencies?filters=NCIT:C3493")
#html <- res$text
#BiocManager::install("GenomicRanges")
#library(jsonlite)
#library(GenomicRanges)

# JSON → R（simplifyVector = FALSE）
#json_data <- fromJSON(html, simplifyVector = FALSE)

#intervals <- json_data$response$results[[1]]$intervalFrequencies

# data.frame
#intervals_df <- as.data.frame(intervals)

# GRanges
#gr <- GRanges(
#  seqnames = intervals_df$referenceName,
#  ranges = IRanges(start = intervals_df$start, end = intervals_df$end),
#  strand = "*",
#  low_gain_frequency = intervals_df$gainFrequency,
#  high_gain_frequency = intervals_df$gainHlfrequency
#)

# GRangesList
#grl <- GRangesList("NCIT:C3493" = gr)

# Result
#grl
```
<br>
```{r}
#Instead, use the rds file shared by Hangjia

freq <- readRDS("./cnv_freq_ncitc3493.rds")
freq
```

To access the CNV frequency data from this object
```{r}
freq[["NCIT:C3493"]]
```
<br>
To get metadata, use `mcols` function:

```{r}
mcols(freq)
```

## Visualise the frequency data
### By genome
```{r, fig.width=12, fig.height=6}
pgxFreqplot(freq)

# Add 40%
axis(side = 2, at = c(-40, 40), labels = c("40", "40"), las = 1, tick = TRUE)

# Add lines at h = heatight
abline(h = 40, lty = 2, col = "grey50", lwd = 1)
abline(h = -40, lty = 2, col = "grey50", lwd = 1)
```
<br>
```{r, fig.width=12, fig.height=3}
chroms <- c(3, 4, 5, 8, 9, 13)

for (chr in chroms) {
  pgxFreqplot(freq, chrom = chr)
  
  # Add 40%
  axis(side = 2, at = c(-40, 40), labels = c("40", "40"), las = 1, tick = TRUE)
  abline(h = 40, lty = 2, col = "grey50", lwd = 1)
  abline(h = -40, lty = 2, col = "grey50", lwd = 1)
}

```

## Analysis
According to the plot, using a threshold 40% for duplication and deletion. 3q, 5p, 8q are frequently duplicated, and 3p, 4p, 4q, 5q, 8p, 9p, 13q are frequently deleted.

## Discussion
Compare to a literature [TCGA LUSC *Nature* 2012](https://www.nature.com/articles/nature11404), they detected frequent gains of 3q (SOX2, PIK3CA, etc.) (70-80%) and 5p (~40%), and loss of 3p (>60%) and 9p21 (CDKN2A) (50-60%) 

* Consistent:

  * Amplifications of 3q, 5p, 8q
  The paper explicitly highlights 3q amplification as highly frequent. While 5p and 8q are not given specific percentages, they are identified as significant GISTIC peaks in Supplementary Figure 2.1b.

  * Deletions of 3p, 9p
  Especially, 9p deletion (CDKN2A locus) is clearly described with a 29% homozygous deletion rate and 72% inactivation overall.

* Unclear: Deletions in 4p, 4q, 5q, 8p, 13q are not explicitly quantified in the paper.

**In conclusion, the result aligns well with the paper for several key regions, especially 3q gain and 9p loss, with supportive overlap for 5p and 8q gains.**
