---
title: "R-exercise_20250410"
author: "Megumi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Basics Exercise Script
Goal: Practice loading data and creating simple visualizations

### 1. Load the Data
Load the `datasets` package (it's built into R, so no need to install it)

```{r}
library(datasets)
```

### 2. Explore the Data
The `iris` dataset is a built-in R dataset with measurements of different flower species

Display the dataset
```{r}
iris 
```
<br>
What is the data type of this dataset?
```{r}
class(iris) 
```
<br>
Display the first 6 rows of the dataset  
```{r}
head(iris)
```
<br>
Get a summary of the dataset (mean, min, max, etc.) 
```{r}
summary(iris)
```
<br>
Get the structure of the dataset (data types, number of rows/columns)  
```{r}
str(iris)
```
<br>
### 3. Assign Variables
Assign petal width and petal length to variables `x` and `y`  
```{r}
x <- iris$Petal.Width
y <- iris$Petal.Length
```
<br>
### 4. Visualization
#### 4a. Create a scatter plot to display the relationship between petal width and petal length
```{r}
plot(x, y, 
     main = "Petal Width vs Petal Length", 
     xlab = "Petal Width", 
     ylab = "Petal Length", 
     col = "blue", 
     pch = 19)
```
<br>
#### 4b. Create a box plot to display Sepal Length for different species  
```{r}
boxplot(Sepal.Length ~ Species, 
        data = iris, 
        main = "Sepal Length by Species", 
        xlab = "Species", 
        ylab = "Sepal Length", 
        col = c("lightgreen", "lightblue", "lightpink"))
```


