---
title: "BIO392-cnv-freq"
author: "Michelle"
date: "2022-09-28"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Install package

```{r}
if (!require(devtools)){ install.packages('devtools')
}
 if (!require(pgxRpi)){ devtools::install_github('progenetix/pgxRpi')
}
library(pgxRpi)
```

## Step2: Search esophageal adenocarcinoma NCIt code

## Step3: Access the CNV frequency data from samples with esophageal adenocarcinoma

```{r}
 freq <- pgxLoader(type='frequency', output='pgxseg',filters='NCIT:C9245', codematches=T)
```

### The retreived data is an object contaning two slots `meta` and `data`.

The `meta` slot looks like this:

```{r}
freq$meta
```

The `data` slot includes two matrices.

```{r}
names(freq$data)
```

The frequency matrix looks like this

```{r}
head(freq$data$`NCIT:C4025`)
```

Dimension of this matrix

```{r}
dim(freq$data$`NCIT:C4025`)
```

## Step4: Visualize data

### By genome

```{r,fig.width=12,fig.height=6}
pgxFreqplot(freq)
```

### By chromosome

```{r,fig.width=12, fig.height=6}
for (x in 1:22){
pgxFreqplot(freq,chrom = x)
}

```

## Step5: Analyse the data

According the plot, we can see frequent gains on chromosome 1q, 3q, 5p, 8q, 16p, 20q
and frequent losses on chromosome 1p, 3p, 8p, 10q, 11q, 16q, 17p, 19p

There is literature where the findings are consistent with 1q,5q,16p,20q-gain and 11q,16q-loss. 
