---
title: "Day4"
author: "Sarah Loose"
date: "2025-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
```{r, warnings = FALSE, Libraries}
library(dplyr)
library(ggplot2)
```

# R Basics Exercise Script
# Goal: Practice loading data and creating simple visualizations

# ---------------------------
# 1. Load the Data
# ---------------------------

# Load the `datasets` package (it's built into R, so no need to install it)
```{r}
library(datasets)
```

 # ---------------------------
# 2. Explore the Data
# ---------------------------

# The `iris` dataset is a built-in R dataset with measurements of different flower species
# Display the dataset 
  
```{r}  
iris 
```

# What is the data type of this dataset?
```{r}
class(iris)

```


# Display the first 6 rows of the dataset  
# HINT: Use the `head()` function
```{r}  
head(iris)
```

 
# Get a summary of the dataset (mean, min, max, etc.)    
```{r}  
summary(iris)
```


# Get the structure of the dataset (data types, number of rows/columns)  
# HINT: Use the `str()` function
```{r}  
str(iris)
```



# ---------------------------
# 3. Assign Variables
# ---------------------------

# Assign petal width and petal length to variables `x` and `y` 
```{r}  
x <- iris$Petal.Width
y <- iris$Petal.Length
```

# ---------------------------
# 4. Visualization
# ---------------------------
## 4a. Create a scatter plot to display the relationship between petal width and petal length     
```{r}    
ggplot(iris, aes(x=x, y =y)) +
  geom_point(color = "skyblue3")+
  geom_smooth(method = lm)
  labs(title = "Petal Width and Length", x = "Petal Width", y = "Petal Length")+
  theme_bw()

```

## 4b. Create a box plot to display Sepal Length for different species      
```{r} 

ggplot(iris, aes(x = Species, y = Sepal.Length, fill = Species)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.5)+
  labs(title = "Sepal Length by Iris Species", x = "Species", y = "Sepal Length")+
  theme_bw()
  
```


```{r}    
# ---------------------------
# Try adding your own comments to explain each step!
# feel free to experiment with other plots or modify colors and labels to make the plot more informative!
```