---
title: "Progenetix Task - Glioblastoma"
output: html_document
date: "2023-09-27"
author: "Roman Stadler"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, cache=F)
```

## R Setup

```{r, eval=F}
install.packages("devtools")
devtools::install_github("progenetix/pgxRpi")
```

```{r}
library(pgxRpi)
```

## Tumor Information

NCIT Code: C3058. 
A glioblastoma is a highly aggressive and malignant brain tumor that originates from glial cells in the central nervous system. 
It is the most common and deadliest form of primary brain cancer in adults, characterized by rapid growth and infiltration into surrounding brain tissue.

## Accessing the Data

```{r}
freq <- pgxLoader(type="frequency", output = "pgxfreq", filters = "NCIT:C3058")
```

```{r}
freq$meta

data <- freq$data
str(data)
```

```{r}
pgxFreqplot(freq)
```


```{r}
pgxFreqplot(freq, chrom=c(7,10,13,18))
``` 

## Analysis

Using a 25% Cutoff, we get:   
- Duplications on 7p, 7q, 19p, 19q, 20p, 20q.  
- Deletions on 9p, 10p, 10q, 13q, 22q.  


[Mirchia et. al.](https://d-nb.info/1202299784/34) claim that "GBM had consistent deletions at 10q23.31".
Furthermore, [Kuga et. al ](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2719013/) describe "losses of chromosomes 9p, 10, 17p, 19q, and 22 and gains at loci on chromosomes 7 and 20."
Therefore, the CNV deletion an duplication pattern is somewhat consistent with the literature, though it does not show a loss that would exceed the 25% margin on the 19q strand.
Striking is the deletion on the chromosome 10, which is visible in the corresponding plot above, which shows exactly what the literature describes as the "consistent deletion".
Moreover, chromosome 13 shows an interesting deletion behaviour, which only extends to the 13q, but not the 13p part of the chromosome.
