---
title: "Day IV - First Steps with R-Markdown"
author: "Sandrin Hunkeler"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: journal
    highlight: tango
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


### 1) Dataset

#### 1.1) Load Dataset
We begin by loading the built-in datasets package, which contains several classic data sets including iris.

```{r }
library(datasets)
```



#### 1.2) Display whole Dataset

The `iris` dataset is a built-in R dataset with measurements of different flower species.


```{r eval=FALSE}
iris 
```


#### 1.3) Display Data Type
What is the data type of the `iris` dataset?

```{r}
class(iris)
```


#### 1.4) Display Data Type of Columns
What is the data type of the `iris` dataset?

```{r}
sapply(iris, class)
```
  
#### 1.5) List First Few Rows
Display the first six rows of the dataset.


```{r}
head(iris)
```


#### 1.6) Summary of Dataset
Get a summary of the dataset (mean, min, max, etc.)  

```{r}
summary(iris)
```


#### 1.7) Structure of Dataset

Get the structure of the dataset (data types, number of rows/columns) 

```{r}
str(iris)
```


### 2) Transformation

#### 2.1) Assign Row Values to Variable

Assign petal width and petal length to variables `x` and `y`  

```{r}
x <- iris$Petal.Width
y <- iris$Petal.Length
```



### 3) Visualization
#### 3.1) Scatter Plot

Create a scatter plot to display the relationship between petal width and petal length  
```{r}
plot(x, y,
     main = "Petal Length vs. Petal Width",
     xlab = "Petal Width (cm)",
     ylab = "Petal Length (cm)",
     pch = 19,
     col = rgb(0.8, 0.1, 0.1, 0.6),
     cex = 1.5,
     xlim = range(x) + c(-0.1, 0.1),
     ylim = range(y) + c(-0.5, 0.5),
     las = 1,
     bty = "l")
grid(col = "gray80", lty = "dotted")

```


#### 3.2) Box Plot
Create a box plot to display Sepal Length for different species  

```{r}
boxplot(Sepal.Length ~ Species,
        data = iris,
        main = "Distribution of Sepal Length by Species",
        xlab = "Species",
        ylab = "Sepal Length (cm)",
        col = c("#66c2a5", "#fc8d62", "#8da0cb"),
        border = "gray40",
        notch = TRUE,
        frame = FALSE,
        las = 1)
grid(nx = NA, ny = NULL, col = "gray90", lty = "dotted")

```



