---
title: "Survival Analysis of Head and Neck Carcinoma"
author: "Lea Scherr & Sandrin Hunkeler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preparations

#### Install packages
```{r eval = FALSE}
install.packages(c("knitr", "dplyr", "survival", "ggplot2", "here", "tibble", "devtools", "ggsurvfit", "gtsummary", "cardx", "survminer", "litedown"))

if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install(version = "3.21")
```



#### Load the libraries
```{r, warning = FALSE, message = FALSE}
library(knitr)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
library(devtools)
library(ggsurvfit)
library(gtsummary)
library(survminer)
library(pgxRpi)
library(GenomicRanges)
library(lubridate)
library(stringr)
```

```{r eval = FALSE}
# Run this code 
devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")
```

## Links
This is where you can obtain the data to [head and neck carcinoma](https://progenetix.org/subset/?id=NCIT:C34447&datasetIds=progenetix) (more info to this on [National cancer institute](https://evsexplore.semantics.cancer.gov/evsexplore/concept/ncit/C34447): [cBioPortal](https://www.cbioportal.org/study/clinicalData?id=ohnca_cptac_gdc)


## Prepare the data

Load the data: 
```{r}
data <- read.delim("./ohnca_cptac_gdc_clinical_data.txt")
```

Filter the data:

For the analysis we need the information from the following columns: Patient ID, Overall Survival (Months), Overall Survival Status, Sex, Alcohol Intensity and Smoking status. 
```{r}
new_data <- select(data, c("Patient.ID", "Overall.Survival..Months.", "Overall.Survival.Status", "Sex", "Alcohol.Intensity", "Smoking.Status"))
```

Wraggle the data:

- The Sex column should be binary for easier use. 
- The Overall.Survival.Status column should also be binary. 
- We want to create Overall.Survival..Days. from Overall.Survival..Months. by multilpying it with 30. 
- The Alcohol Abuse should also be in numbers (factors for later use).
- The Smoking data should also be in numbers (factors for later use).
```{r}
new_data <- mutate(new_data, Sex = recode(Sex, `Male` = 0, `Female` = 1))

new_data <- mutate(new_data, 
                   Overall.Survival.Status = recode(Overall.Survival.Status, 
                                                              `1:DECEASED` = 1, 
                                                              `0:LIVING` = 0))

new_data <- mutate(new_data, Overall.Survival..Days. = Overall.Survival..Months. * 30)

new_data <- mutate(new_data, Alcohol.Abuse = recode(Alcohol.Intensity, 
                                                    'Occasional Drinker' = 2, 
                                                    'Non-Drinker' = 1, 
                                                    'Lifelong Non-Drinker' = 0))
new_data$Alcohol.Abuse <- as.factor(new_data$Alcohol.Abuse)

new_data <- mutate(new_data, Smoking = recode(Smoking.Status, 
                                    "Lifelong Non-Smoker" = 0, 
                                    "Current Reformed Smoker for < or = 15 yrs" = 1, 
                                    "Current Reformed Smoker for > 15 yrs" = 2, 
                                    "Current Smoker" = 3, 
                                    "Smoking history not documented" = NaN,
                                    "Current Reformed Smoker, Duration Not Specified" = NaN,   # since this cannot be assigned, they will not be used for the analysis because we suspect a difference between reformed less than 15 years and more than 15 years
                                    ))
new_data$Smoking <- as.factor(new_data$Smoking)
```


### Getting to know the data

How many Samples are there in the dataset from males and females.
```{r}
ggplot(data=new_data, aes(x=Sex)) +
  geom_bar(aes(fill = factor(Sex, labels = c("Male (0)", "Female (1)")))) +
  scale_fill_manual(
    values = c("skyblue", "lightpink"),
    name = "Sex",
    labels = c("Male (0)", "Female (1)")
  ) +
  xlab("Sex (0 = Male, 1 = Female)") +
  ylab("Samples count") +
  labs(caption = "Note: Only values 0 and 1 appear because 'Sex' is a binary variable (no intermediate values).") +
  theme_minimal()

table(new_data$Sex)
```


```{r}
ggplot(data=new_data, aes(x=Alcohol.Intensity)) +
  geom_bar() +
  xlab("Alcohol consumption") +
  ylab("Samples count") +
  theme_minimal()
```

```{r}
ggplot(data = new_data, aes(x = factor(Smoking, levels = c(0, 1, 2, 3), 
                            labels = c("Lifelong Non-Smoker (0)",
                                       "Reformed ≤15 yrs (1)",
                                       "Reformed >15 yrs (2)",
                                       "Current Smoker (3)")))) +
  geom_bar() +
  xlab("Tobacco Consumption") +
  ylab("Sample Count") +
  theme_minimal() +
  labs(caption = "Note: 'Smoking history not documented' and NA values are excluded")
```

## Kaplan-Meier plot

The explicit formulas for a Kaplan-Meier and Cox plot can be found from this [Survival Analysis by Ramaa Nathan](https://ramaanathan.github.io/SurvivalAnalysis/#223_semi-parametric_cox_proportional_hazard_regression_model).

### Model the data
Use the code from [Emily C. Zabor](https://www.emilyzabor.com/survival-analysis-in-r.html#the-cox-regression-model) to fit the survival days to the survival status.
```{r}
s1 <- survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, 
              data = new_data)
str(s1)
```

### Plot the survival data
This code gives the survival plot. 
```{r}
survfit2(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, 
         data = new_data) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +
  add_confidence_interval() +
  add_risktable()
```

This code gives the Kaplan-Meier plot with and without censoring. 
```{r}
ggsurvplot(survfit(Surv(Overall.Survival..Days.) ~ Overall.Survival.Status, data = new_data),
           data = new_data,
           conf.int = TRUE,                         # Confidence intervals
           pval = FALSE,                            # Add log-rank test p-value if needed
           risk.table = TRUE,                       # Add risk table below plot
           legend.title = "Group",
           legend.labs = c("Ignoring censoring", "With censoring"),
           palette = c("#2E9FDF", "#E7B800"),        # Custom colors
           xlab = "Days",
           ylab = "Overall survival probability",
           ggtheme = theme_minimal())
```

### Yearly survival probabilities
1-year survival
```{r}
summary(survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, data = new_data), times = 365.25)
```

2-year survival
```{r}
summary(survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, data = new_data), times = 730.5)
```

3-year survival
```{r}
summary(survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, data = new_data), times = 1095.75)
```

4-year survival
```{r}
summary(survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, data = new_data), times = 1461)
```


### Few calculations
Statistically calculated mean. This mean is faulty and should not be used.  
```{r}
new_data %>% 
  filter(Overall.Survival.Status == 1) %>% 
  summarize(median_surv = median(Overall.Survival..Days.))
```

Fitted mean with confidence intervals. This mean can be used. 
```{r eval = FALSE}
tbl_survfit(survfit(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ 1, data = new_data), 
  probs = 0.5,
  label_header = "**Median survival (95% CI)**"
  )
```


### Comparison between groups
#### Sex
Fitting the Sex data and obtaining Chisq and p-value. 
```{r}
survdiff(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Sex, data = new_data)
```

#### Alcohol Abuse
Fitting the Alcohol abuse data.
```{r}
survdiff(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Alcohol.Abuse, data = new_data)
```

#### Smoking
Fitting the Smoking data.
```{r}
survdiff(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking, data = new_data)
```

#### Smoking + Alcohol Abuse
Fitting the interaction between Smoking and Alcohol abuse.
```{r}
survdiff(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking + Alcohol.Abuse, data = new_data)
```


## Cox Regression
Done with the codes from [Statistical tools for high-throughput data analysis](https://www.sthda.com/english/wiki/cox-proportional-hazards-model) and [Emily C. Zabor](https://www.emilyzabor.com/survival-analysis-in-r.html#the-cox-regression-model).


### Survival in relation to Sex

Fitting it with the Cox-function to obtain the Cox p-value.
```{r}
sexes <- coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Sex, data = new_data)
sexes
```

Create a table of the data with the Hazard ratio, confidence intervals and p-values.
```{r eval = FALSE}
coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Sex, data = new_data) %>% 
  tbl_regression(exp = TRUE) 
```

Create a plot for the survival probability in relation to sex.
```{r}
newdata1 <- expand.grid(Sex = factor(c(0,1), levels = c(0,1)))
ggsurvplot(survfit(sexes, newdata = newdata1),
           palette = c("blue", "red"),
           data = new_data,
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Male", "Female"),
           ggtheme = theme_minimal())
```


### Survival in relation to Alcohol abuse

Fitting it with the Cox-function to obtain the Cox-fitted p-value. 
```{r}
alcohol <- coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Alcohol.Abuse, data = new_data)
alcohol
```

Create a table of the data with the Hazard ratio, confidence intervals and p-values.
```{r eval = FALSE}
coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Alcohol.Abuse, data = new_data) %>% 
  tbl_regression(exp = TRUE) 
```

Create a plot for the survival probability in relation to Alcohol abuse.
```{r}
newdata2 <- expand.grid(Alcohol.Abuse = factor(c(0,1,2), levels = c(0,1,2)))
ggsurvplot(survfit(alcohol, newdata = newdata2),
           palette = c("#2E9FDF", "#E7B800", "#FC4E07"),
           data = new_data,
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Lifelong non-drinker", "Non-drinker", "Occasional drinker"),
           ggtheme = theme_minimal())
```



### Survival in relation to Smoking

Fitting it with the Cox-function to obtain the Cox-fitted p-value.
```{r}
smoke <- coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking, data = new_data)
smoke
```

Create a table of the data with the Hazard ratio, confidence intervals and p-values.
```{r eval = FALSE}
coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking, data = new_data) %>% 
  tbl_regression(exp = TRUE) 
```

Create a plot for the survival probability in relation to Smoking.
```{r}
newdata3 <- expand.grid(Smoking = factor(c(0,1,2,3), levels = c(0,1,2,3)))
ggsurvplot(survfit(smoke, newdata = newdata3),
           palette = c("#2E9FDF", "#E7B800", "#FC4E07", "green"),
           data = new_data,
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Lifelong non-smoker", "Current Reformed Smoker for < or = 15 yrs", "Current Reformed Smoker for > 15 yrs", "Current Smoker"),
           ggtheme = theme_minimal())
```



### Survival in relation to the interaction of Smoking and Alcohol abuse

Fitting it with the Cox-function to obtain the Cox-fitted p-value.
```{r}
smo_alk <- coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking + Alcohol.Abuse, 
                 data = new_data)
smo_alk
```

Create a table of the data with the Hazard ratio, confidence intervals and p-values.
```{r eval = FALSE}
coxph(Surv(Overall.Survival..Days., Overall.Survival.Status) ~ Smoking + Alcohol.Abuse, data = new_data) %>% 
  tbl_regression(exp = TRUE) 
```

We decided not to make a plot for the this interaction, because when we did, it did not create a useful output. 


## Comparison with progenetix database

Get the data from the [progenetix](https://progenetix.org/) database.
```{r}
freq <- pgxLoader(type = "individuals", filters = "NCIT:C34447")
head(freq)
```

Create a new dataset with the relevant columns:
- The individual ID
- The sex of the individual
- Days to death/event
- Status of the individual (alive/dead)
```{r}
pgx_data <- select(freq, "individual_id", "sex_label", "info_days_to_death", "info_death")
```

Wraggle the data:
```{r}
pgx_data <- mutate(pgx_data, Sex = recode(sex_label, "male" = 0, "female" = 1))

pgx_data <- mutate(pgx_data, Status = recode(info_death, "alive" = 0, "dead" = 1))

pgx_data <- mutate(pgx_data, Days.to.Death = as.numeric(info_days_to_death))
```

Fit the data using the `survfit` function.  
```{r}
s2 <- survfit(Surv(Days.to.Death, Status) ~ 1, 
              data = pgx_data)
str(s2)
```

### Plot the survival data
This code gives the survival plot. 
```{r}
survfit2(Surv(Days.to.Death, Status) ~ 1, 
         data = pgx_data) %>% 
  ggsurvfit() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  ) +
  add_confidence_interval() +
  add_risktable()
```

```{r}
tbl_survfit(survfit(Surv(Days.to.Death, Status) ~ 1, data = pgx_data), 
  probs = 0.5,
  label_header = "**Median survival (95% CI)**"
  )
```

### Comparison between groups
#### Sex
Fitting the Sex data and obtaining Chisq and p-value. 
```{r}
survdiff(Surv(Days.to.Death, Status) ~ Sex, data = pgx_data)
```

## Cox Regression

### Survival in relation to Sex

Fitting it with the Cox-function to obtain the Cox p-value.
```{r}
sexes_cox <- coxph(Surv(Days.to.Death, Status) ~ Sex, data = pgx_data)
sexes_cox
```

Create a table of the data with the Hazard ratio, confidence intervals and p-values.
```{r}
coxph(Surv(Days.to.Death, Status) ~ Sex, data = pgx_data) %>% 
  tbl_regression(exp = TRUE) 
```

Create a plot for the survival probability in relation to sex.
```{r}
newdata5 <- expand.grid(Sex = factor(c(0,1), levels = c(0,1)))
ggsurvplot(survfit(sexes_cox, newdata = newdata5),
           palette = c("blue", "red"),
           data = pgx_data,
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Male", "Female"),
           ggtheme = theme_minimal())
```

Scaled to our plot:
```{r}
newdata5 <- expand.grid(Sex = factor(c(0,1), levels = c(0,1)))
ggsurvplot(survfit(sexes_cox, newdata = newdata5),
           palette = c("blue", "red"),
           data = pgx_data,
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Male", "Female"),
           ggtheme = theme_minimal(),
           xlim= c(0, 2000))
```


