---
title: "cnv-freq-analysis-template"
author: "Hangjia Zhao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages

```{r eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.20")

BiocManager::install(c("pgxRpi", "GenomicRanges"))
```

## Load packages

```{r}
library(pgxRpi)
library(GenomicRanges)
```

## Disease description

I'd like to study esophageal adenocarcinoma. In Progenetix, the corresponding NCIt code is "NCIT:C4025". ...

## Access CNV frequency data from samples with esophageal adenocarcinoma

```{r}
freq <- pgxLoader(type="cnv_frequency",filters="NCIT:C4025")
freq
```

To access the CNV frequency data from this object

```{r}
freq[["NCIT:C4025"]]
```

To get metadata, use `mcols` function:

```{r}
mcols(freq)
```
## Visualize the frequency data

### By genome

```{r, fig.width=12, fig.height=6}
pgxFreqplot(freq)
```

### By chromosome

```{r, fig.width=12, fig.height=6}
pgxFreqplot(freq,chrom = c(8, 20), layout = c(2,1))
```

## Analysis

According to the plot, using a threshold 25% for duplication and deletion. 7p, 8q, 20p, 20q are frequently duplicated, 4p, 4q, 5q, 9p, 17p, 18q, 21q are frequently deleted. 

Comapred to another literature ([link](https://www.nature.com/articles/modpathol2008150)), they detected frequent gains of 6p, 8q, 13q, 17q, 20q, and losses of 4q, 5q, 15q, and 18q. Same? Different? Why? 




