---
title: "cnv_freqBreastCancer.Rmd"
date: "2022-09-28"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

 ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Install package

```{r}
if (!require(devtools)){ install.packages('devtools')
}
 if (!require(pgxRpi)){ devtools::install_github('progenetix/pgxRpi')
 }
library(pgxRpi)
```

## Step2: Search esophageal adenocarcinoma NCIt code

## Step3: Access the CNV frequency data from samples with esophageal adenocarcinoma

```{r}
 freq <- pgxLoader(type='frequency', output='pgxseg',filters='NCIT:C9245',
                  codematches=T)



```

### The retreived data is an object contaning two slots `meta` and `data`.

The `meta` slot looks like this:
  
```{r}
freq$meta

```

The `data` slot includes two matrices.

```{r}
names(freq$data)
```

The frequency matrix looks like this

```{r}
head(freq$data$`NCIT:C9245`)
```

Dimension of this matrix

```{r}
dim(freq$data$`NCIT:C9245`)
```

## Step4: Visualize data

### By genome

```{r,fig.width=12,fig.height=6}
pgxFreqplot(freq)
```

### By chromosome

```{r,fig.width=12, fig.height=6}
for (y in 1:22) {
  pgxFreqplot(freq,chrom = y)
}

```

## Step5: Analyse the data

According the plot, we can see frequenct gains on chromosome 1q, 3q, 8q and 20q.
and frequenct losses on chromosome 1p, 3p, 8p, 9p, 9q, 10q, 11q, 16q, 17p and 19p.


A more detailed use case see this [link](https://htmlpreview.github.io/?https://github.com/progenetix/pgxRpi/blob/main/vignettes/Introduction_2_loadfrequency