---
title: "BIO392-Patricia-Freitag-CNV-freq"
author: "Patricia Freitag"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tumor information 

Glioblastomas are a highly malignant form of brain cancer. They originate from abnormal growth of glial cells. 
The NCIt Code is: C4025

# Access the CNV frequency data

## load library

```{r}

library(pgxRpi)

```

## access data 

```{r}
freq <- pgxLoader(type = "frequency", output = "pgxfreq", filters = "NCIT:C3058")
```

The metadata information is

```{r}
freq$meta
```

The data looks like this

```{r}
freq$data$"NCIT:C3058"[c(1:3),]
```

## Visualization

```{r}
pgxFreqplot(freq)
```

According to plot, there are frequent duplications on chromosome 7p,7q,19p,19q,20p and 20q and deletions on chromosomes 10p, 10q and 13q. The cutoff percentage choosen is 25%. 
One [study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8108987/) detected CNA gains which were mainly located on Chromosome 7p and 7q and CNA losses located on Chromosome 10q. There are overlaps between these two finding: Gains on 7p and 7q, loss on 10q
