---
title: "D04 R Exercise"
author: "BaBurri"
date: "2023-05-07"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load the Data

```{r , include=TRUE}
library("datasets")

```

## 2. Explore the Data

#### Display the dataset
```{r , include=TRUE}
# Combine the first 3 and last 3 rows of the iris dataset
iris_subset <- rbind(head(iris, 3), tail(iris, 3))

# Display the result
print(iris_subset)
```

#### Data type of this dataset
```{r , include=TRUE}
class(iris)
```

#### Display the first 6 rows of the dataset
```{r , include=TRUE}
head(iris)
```

#### Summary of the dataset (mean, min, max, etc.)
```{r , include=TRUE}
summary(iris)
```

#### Structure of the dataset (data types, number of rows/columns) 
```{r , include=TRUE}
str(iris)
```

## 3. Assign Variables

#### Assign petal width and petal length to variables "x" and "y"  
```{r 3. Assign Variables, include=TRUE}
x <- iris$Petal.Width
y <- iris$Petal.Length
```

## 4. Visualization

#### Scatter-plot to display the relationship between petal width and petal length


```{r , fig.width=5, fig.height=5, include=TRUE}
plot(iris$Petal.Length, iris$Petal.Width, pch = 20,
     main = "Petal Length vs. Petal Width",
     xlab = "Petal Length",
     ylab = "Petal Width")

```

#### Box-plot to display Sepal Length for different species  

```{r , fig.width=5, fig.height=5, include=TRUE}
boxplot(Sepal.Length ~ Species, data = iris,
        main = "Sepal Length by Species",
        xlab = "Species",
        ylab = "Sepal Length")

```
