---
title: "Project 1"
author: "Sara Burckhardt and Majorie Baertschi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load data  

```{r}
# Load required library
library(ggplot2)

# Read the eigenvec file
pca <- read.table("chr3_pca.eigenvec", header = FALSE)

# Add column names
colnames(pca) <- c("FID", "sample", paste0("PC", 1:(ncol(pca) - 2)))


```
Bring the IID and the corresponding subpopulation/ population together 

```{r}
pop <- read.table("sample_type.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)


#add the population and superpopulation to the pca table 

result <- merge(pca, pop, by = "sample", all.x = TRUE)


```


## PCA Plots
Plot of the PC1 and PC2 colored by Super-population and Population:

```{r}
ggplot(result, aes(x = PC1, y = PC2, color = as.factor(Superpopulation))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Super-population", x = "Principal Component 1", y = "Principal Component 2") +
  theme_classic()

```
```{r}
ggplot(result, aes(x = PC1, y = PC2, color = as.factor(Population))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Population", x = "Principal Component 1", y = "Principal Component 2") +
  theme_classic()
```

Plots of the PC3-10 
```{r echo=FALSE}
ggplot(result, aes(x = PC3, y = PC4, color = as.factor(Superpopulation))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Super-Population", x = "Principal Component 3", y = "Principal Component 4") +
  theme_classic()


ggplot(result, aes(x = PC5, y = PC6, color = as.factor(Superpopulation))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Super-Population", x = "Principal Component 5", y = "Principal Component 6") +
  theme_classic()

ggplot(result, aes(x = PC7, y = PC8, color = as.factor(Superpopulation))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Super-Population", x = "Principal Component 7", y = "Principal Component 8") +
  theme_classic()

ggplot(result, aes(x = PC9, y = PC10, color = as.factor(Superpopulation))) +
  geom_point(size = 2, alpha = 0.8) +
  labs(title = "PCA by Super-Population", x = "Principal Component 9", y = "Principal Component 10") +
  theme_classic()
```

The higher the PC number the less important is the variance, so the most meaningful are the PC1 and PC2 which explains the first and second most variance. 


##Admixture Plots

Import data 
```{r}
# Read Q file
q_matrix <- read.table("chr3_for_admixture_plink1.5.Q")

# Read sample IDs from FAM file
fam <- read.table("chr3_for_admixture_plink1.fam")
sample_ids <- fam$V2  # 2nd column is usually individual ID

# Add sample IDs to Q matrix
q_matrix$Sample <- sample_ids

```


Due to long format reshape the data for plotting
```{r}
library(tidyr)
library(dplyr)

# Rename columns for clarity
K <- ncol(q_matrix) - 1
colnames(q_matrix)[1:K] <- paste0("Cluster", 1:K)

# Convert to long format
q_long <- q_matrix %>%
  pivot_longer(cols = starts_with("Cluster"), names_to = "Cluster", values_to = "Proportion")

```

Make the Plots
```{r}
library(ggplot2)

ggplot(q_long, aes(x = Sample, y = Proportion, fill = Cluster)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  theme(axis.text.x = element_blank(),  # Hide sample names if too many
        axis.ticks.x = element_blank()) +
  labs(x = "Individuals", y = "Ancestry Proportion", title = "ADMIXTURE Plot") +
  scale_fill_brewer(palette = "Set2")  # or use your preferred color palette

```


Bring the Sample and the corresponding subpopulation/ population together 

```{r}

#add the population and superpopulation to the pca table 
colnames(q_long)[1] <- "sample"

results_2 <- merge(q_long, pop, by = "sample", all.x = TRUE)
```


```{r}
ggplot(results_2, aes(x = Superpopulation, y = Proportion, fill = Cluster)) +
  geom_bar(stat = "identity") +
  theme_bw() +
  #theme(axis.text.x = element_blank(),  # Hide sample names if too many
       # axis.ticks.x = element_blank()) +
  labs(x = "Superpopulation", y = "Ancestry Proportion", title = "ADMIXTURE Plot") +
  scale_fill_brewer(palette = "Set2")  # or use your preferred color palette
```

```{r}
#sort by superpopulations
library(dplyr)

results_2 <- results_2 %>%
  arrange(Superpopulation, sample) %>%
  mutate(sample = factor(sample, levels = unique(sample)))

#make a faced plot
ggplot(results_2, aes(x = sample, y = Proportion, fill = Cluster)) +
  geom_bar(stat = "identity", width = 1) +
  facet_grid(~ Superpopulation, scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    strip.text = element_text(size = 8)
  ) +
  labs(x = "Individuals", y = "Ancestry Proportion", title = "ADMIXTURE Plot by Population")


```
```{r fig.width=10, fig.height=4}
results_2 <- results_2 %>%
  arrange(Population, sample) %>%
  mutate(sample = factor(sample, levels = unique(sample)))


#make a faced plot
ggplot(results_2, aes(x = sample, y = Proportion, fill = Cluster)) +
  geom_bar(stat = "identity", width = 1) +
  facet_grid(~ Population, scales = "free_x", space = "free_x") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    strip.text = element_text(size = 8)
  ) +
  labs(x = "Individuals", y = "Ancestry Proportion", title = "ADMIXTURE Plot by Population")
```

