---
title: "CNV-freq-Glioblastoma"
author: "Tina Sidler"
date: "10/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Install package

```{r}
if (!require(devtools)){
    install.packages('devtools')
}
devtools::install_github('progenetix/pgxRpi')
library(pgxRpi)
```

## Step2: Search esophageal adenocarcinoma NCIt code


## Step3: Access the CNV frequency data from samples with esophageal adenocarcinoma

```{r}
freq <- pgxLoader(type='frequency', output='pgxseg',filters='NCIT:C3058',
                  codematches=T)
```

### The retreived data is an object contaning two slots `meta` and `data`.

The `meta` slot looks like this:

```{r}
freq$meta
```

The `data` slot includes two matrices.

```{r}
names(freq$data)
```

In this case, these two matrices are the same.

```{r}
head(freq$data$`NCIT:C3058`)
```

```{r}
dim(freq$data$`NCIT:C3058`)
```

## Step4: Visualize data

### By genome

```{r}
pgxFreqplot(freq)
```

### By chromosome

```{r}
pgxFreqplot(freq,chrom = 22)
```

## Step5: Analyse the data

According the plot, we can see frequenct gains on chromosome 7p, 7q, 19p, 19q 20p,20q 
and frequenct losses on chromosome 9p, 10p, 10q, 13q, 22q.

There is a literature where the findings are consistent with the majority of mine. Here is the paper-link: https://www.sciencedirect.com/science/article/pii/S0046817718304970