---
title: "bio392-cnv-freq"
author: "Kasia Speichert"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tumor information

Pancreatic ductal adenocarcinoman (PDAC) is a highly devastating disease which accounts for more than 90% of all pancreatic malignancies. PDAC is the fourth most frequent cause of cancer-related deaths. The overall 5-year survival rate accounts to less than 8%. 

https://ro-journal.biomedcentral.com/articles/10.1186/s13014-019-1345-6


The NCIt code is Code C9120

#Access the CNV frequency data 

## Load library 

```{r}
library(pgxRpi)
```

## Access data 

```{r}
freq <- pgxLoader(type='frequency',output = 'pgxfreq',filters = 'NCIT:C9120')
```

The metadata information is

```{r}
freq$meta
```
The data looks like this

```{r}
freq$data$`NCIT:C9120`[c(1:3),]
```
```{r}
pgxFreqplot(freq, chrom = 7)
```



#Visualization
```{r}
pgxFreqplot(freq)
```


```{r}
pgxFreqplot(freq, chrom = 1)
```

```{r}
pgxFreqplot(freq, chrom = 1)
```

## Analysis

According to the plot, there are frequent gains on chromosome 1q, 2, 3q, 5, 7, 8q, 11q, 12q, 14, 15, 16, 17q, 20q and frequent deletion on chromosomes 1p, 3p, 4, 6, 8p, 9p, 13q, 17p, 18q, 21, 22.

As a comparison the data from the paper [https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2492386/] showed pretty similar results with some minor differences:
“We detected frequent gains of 1q, 2, 3, 5, 7p, 8q, 11, 14q and 17q (≥78% of cases), and losses of 1p, 3p, 6, 9p, 13q, 14q, 17p and 18q (≥44%)“.