---
title: "Survival analysis in Bladder Cancer"
author: "Sarah Oppolzer and Luisa Buchegger"
date: "`r Sys.Date()`"
output: html_document
---

Blockkurs Project

Lets install Packages etc. 
```{r}
# Install packages if needed
rm(list=ls())
install.packages(c("knitr", "dplyr", "survival", "ggplot2", "here", "tibble"))
library(knitr)
library(dplyr)
library(survival)
library(ggplot2)
library(tibble)
install.packages("devtools")
devtools::install_github("zabore/ezfun")
ezfun::set_ccf_palette("contrast")

install.packages(c("lubridate", "ggsurvfit", "gtsummary", "tidycmprsk"))
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)

devtools::install_github("zabore/condsurv")
library(condsurv)
library(tidyverse)
install.packages("cardx")
library(survival)
library(gtsummary)
library(survminer)
```

Take the data from: https://www.cbioportal.org/study/clinicalData?id=blca_tcga_gdc (Bladder Urothelial Carcinoma )
then follow tutorial: https://www.emilyzabor.com/survival-analysis-in-r.html
```{r}
dd <- read.delim("~/Desktop/MSc/FS_25/Molecular_Sequence_Variations/Presentation/blca_tcga_gdc_clinical_data.tsv") 
colnames(dd)
str(dd)
```

Select the relevant colums and encode categorical variables (dead: 0, alive: 1, mann: 0, frau: 1)
ALso, we change Months to days!
```{r}
dd_rel <- select(dd,sampleId,VITAL_STATUS,OS_MONTHS,SEX,AGE,SMOKING_PACK_YEARS, MUTATION_COUNT)

dd<- dd%>%
  mutate(VITAL_STATUS = ifelse(VITAL_STATUS == "Dead", 1, 0))

dd_rel <- dd_rel%>%
  mutate(VITAL_STATUS = ifelse(VITAL_STATUS == "Dead", 1, 0))

dd_rel <- dd_rel%>%
  mutate(SEX = ifelse(SEX == "Male", 0, 1))

dd_rel <- dd_rel %>%
  mutate(SMOKING_PACK_YEARS = ifelse(is.na(SMOKING_PACK_YEARS), 0, 1))

dd_rel$OS_DAYS <- dd_rel$OS_MONTHS * 30.44
```

KAPLAN MEIER SURVIVAL CURVE
lets make the Kaplan Meier survival curve
```{r}
Surv(dd_rel$OS_MONTHS, dd_rel$VITAL_STATUS)[1:10]

#install.packages("ggsurvfit")
s1 <- survfit(Surv(OS_MONTHS, VITAL_STATUS) ~ 1, data = dd_rel)
str(s1)

survfit2(Surv(OS_MONTHS, VITAL_STATUS) ~ 1, data = dd_rel) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability")+
  add_confidence_interval()+
  add_risktable()
```

Estimating x-year survival -> for one year survival probability
```{r}
summary(survfit(Surv(OS_MONTHS, VITAL_STATUS)  ~ 1, data = dd_rel), times =12)
```

Table for 1 year survival
```{r}
survfit(Surv(OS_MONTHS, VITAL_STATUS) ~ 1, data = dd_rel) %>% 
  tbl_survfit(
    times = 12,
    label_header = "**1-year survival (95% CI)**"
  )
```


Median survival
```{r}
survfit(Surv(OS_MONTHS, VITAL_STATUS) ~ 1, data = dd_rel) %>% 
  tbl_survfit(
    probs = 0.5,
    label_header = "**Median survival (95% CI)**"
  )
```

Comparing survival times between sexes
```{r}
survdiff(Surv(OS_MONTHS, VITAL_STATUS) ~ SEX, data = dd_rel)
```
No significant difference between women and men!

COX PROPORTIONAL HAZARDS MODEL 
Basic COX regression model
```{r}

# Fit model with a group variable
cox_model <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ 1, data = dd_rel)
 
# Create survival curves
surv_fit <- survfit(cox_model)
 
# Plot
ggsurvplot(surv_fit, data = dd_rel, conf.int = TRUE)
#ggsurvplot(surv_fit, data = filtered, conf.int = TRUE, risk.table = TRUE)
 
```

COX Regression for sex
```{r}
coxph(Surv(OS_MONTHS, VITAL_STATUS) ~ SEX, data = dd_rel)
```
HR: higher then 1: first group higher risk that the event happens. for each year: 4%

COX Model for age
```{r}

cox_model <- coxph(Surv(OS_MONTHS, VITAL_STATUS) ~ AGE, data = dd_rel) 

```

MIltivariable COX Regression Model - For 2 age groups and Sex
```{r}
cox_model2 <- coxph(Surv(OS_MONTHS, VITAL_STATUS) ~ AGE+SEX, data = dd_rel)


# Create a new data frame with hypothetical values or groupings
newdata <- expand.grid(
  AGE = c(40, 70),
  SEX = c("0", "1"))  # Assuming "0" = female, "1" = male
# Predict and plot
ggsurvplot(survfit(cox_model2, newdata = newdata), 
           data = dd_rel, 
           conf.int = TRUE,
           legend.title = "Profile",
           legend.labs = c("Female, 40", "Female, 70", "Male,40", "Male, 70"),
           xlab = "Time (months)",
           ylab = "Survival probability")
```
Now with age classes
```{r}
dd_rel$OS_DAYS <- dd_rel$OS_MONTHS * 30.44
dd_rel$AGE_GROUP <- cut(dd_rel$AGE,
                        breaks = c(0, 60, 100), 
                        labels = c("0-60", "60-100"),
                        right = FALSE)

cox_model3 <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ AGE_GROUP, data = dd_rel)

newdata2 <- expand.grid(
  AGE_GROUP = c("0-60", "60-100"))

ggsurvplot(
  survfit(cox_model3, newdata = newdata2),
  data = dd_rel,
  conf.int = TRUE,
  legend.title = "Profile",
  legend.labs = c("0-60", "60-100"),
  xlab = "Time (days)",
  ylab = "Survival probability"
)
```
Look at Ages distribution 

```{r}
library(ggplot2)
ggplot(dd, aes(x=AGE))+
  geom_histogram()
```
```{r}
survdiff(Surv(OS_MONTHS, VITAL_STATUS) ~ AGE_GROUP, data = dd_rel)
```
Survival rate between young and older people is significant. 

SMOKING COX REGRESSION
```{r}
dd_rel$OS_DAYS <- dd_rel$OS_MONTHS * 30.44

cox_model4 <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ SMOKING_PACK_YEARS, data = dd_rel)

newdata4 <- data.frame(SMOKING_PACK_YEARS = c(0, 1))

ggsurvplot(
  survfit(cox_model4, newdata = newdata4),
  data = dd_rel,
  conf.int = TRUE,
  legend.title = "Smoking Status",
  legend.labs = c("Non-smoker", "Smoker"),
  xlab = "Time (days)",
  ylab = "Survival probability"
)
```
```{r}
survdiff(Surv(OS_MONTHS, VITAL_STATUS) ~ SMOKING_PACK_YEARS, data = dd_rel)
```
We notice somethign werid: in our data, smoking does not seem to be a risk facto for bladder cancer?? Literature says differently

Lets try classes
```{r}
summary(dd$SMOKING_PACK_YEARS)

dd$OS_DAYS <- dd$OS_MONTHS * 30.44

dd$SMOKING_CUT<- cut(dd$SMOKING_PACK_YEARS,
                      breaks = c(0, 30, 100),  # Defining the ranges
                      labels = c("0-30", "30-100"),  # Naming the categories
                      right = FALSE)  # Including 0 but not 25 in the first category

cox_model5 <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ SMOKING_CUT, data = dd)

# Prepare newdata for plotting
newdata5 <- expand.grid(SMOKING_CUT = c("0-30", "30-100"))

# Plot
ggsurvplot(
  survfit(cox_model5, newdata = newdata5),
  data = dd,
  conf.int = TRUE,
  legend.title = "Smoking Category",
  legend.labs = c("0-30 pack years", "30-100 pack years"),
  xlab = "Time (days)",
  ylab = "Survival probability"
)
```
signifikant? (viel rauchen)
```{r}
survdiff(Surv(OS_MONTHS, VITAL_STATUS) ~ SMOKING_CUT, data = dd)
```
```{r}
ggplot(dd, aes(x= OS_MONTHS, y= SMOKING_PACK_YEARS))+
  geom_point()
```
Still no significant difference: no reliable smoking data in our data set (manny missing values that might be smokers too)

MUTATIONAL LOAD
```{r}
dd_rel$OS_DAYS <- dd_rel$OS_MONTHS * 30.44
dd_rel$MUTATION_COUNT <- cut(dd_rel$MUTATION_COUNT,
                        breaks = c(0, 144, 1500), 
                        labels = c("0-144", "144-1500"),
                        right = FALSE)

cox_model6 <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ MUTATION_COUNT, data = dd_rel)

newdata6 <- expand.grid(
  MUTATION_COUNT = c("0-144", "144-1500"))

ggsurvplot(
  survfit(cox_model6, newdata = newdata6),
  data = dd_rel,
  conf.int = TRUE,
  legend.title = "Profile",
  legend.labs = c("0-144", "144-1500"),
  xlab = "Time (days)",
  ylab = "Survival probability"
)
```
Distribution Mutational Load
```{r}
ggplot(dd, aes(x=MUTATION_COUNT))+
  geom_histogram()
summary(dd$MUTATION_COUNT)
```
```{r}
survdiff(Surv(OS_MONTHS, VITAL_STATUS) ~ MUTATION_COUNT, data = dd_rel)
```
Significant difference in Survival for different mutational loads

Lets look at stages
```{r}
unique(dd$PATH_STAGE)
library(dplyr)

dd <- dd %>%
  filter(!is.na(PATH_STAGE) & PATH_STAGE != "" & PATH_STAGE != "NA")
```

Lets make a COX analysis of that
```{r}
dd$OS_DAYS <- dd$OS_MONTHS * 30.44

cox_model7 <- coxph(Surv(OS_DAYS, VITAL_STATUS) ~ PATH_STAGE, data = dd)

newdata7 <- data.frame(PATH_STAGE = c("Stage I", "Stage II", "Stage III", "Stage IV"))


ggsurvplot(
  survfit(cox_model7, newdata = newdata7),
  data = dd,
  conf.int = TRUE,
  legend.title = "Stage Number:",
  legend.labs = c("Stage I", "Stage II", "Stage III", "Stage IV"),
  xlab = "Time (days)",
  ylab = "Survival probability"
)
```
Lets try it in a Kaplan Meyer curve
```{r}
survfit2(Surv(OS_MONTHS, VITAL_STATUS) ~ PATH_STAGE, data = dd) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability"
  ) +
  add_confidence_interval() +
  add_risktable()

```
We see, that stage is a good predictor for risk factor. Interesting difference in Kaplan Meier and Cox: Cox predicts values, Kaplan jsut shows the ones that exist (see blue and yellow line in the Kaplan Meier Plot ending early!)
