---
title: "CNV frequencies in tumor types"
author: "Maeve Lin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Tumor information
Ependymoma is a type of brain tumor that mostly occurs in children and young adults. It is slow growing and the most common ependymal neoplasm. It starts in the ependymal cells lining the brain ventricles and the spinal cord central canal. 


the NCIt code is C3017

# Access the CNV frequency data

## Load library

```{r}
library(pgxRpi)
```


## Access data
```{r}
freq <- pgxLoader(type = 'frequency', output = 'pgxfreq', filters = "NCIT:C3017") # add a second filter if you want to get data of two tumors
```

the metadata information is
```{r}
freq$meta
```
progenetix has 537 samples for this tumor


the data looks like this
```{r}
freq$data$`NCIT:C3017`[c(1:3),]
```


# Visualization
```{r}
pgxFreqplot(freq)
```
have a closer look at chr 9, chr 22, chr 6, chr 7
I choose a cutoff value of 20%. 


```{r}
pgxFreqplot(freq, chrom = 1)
```
1q duplication slightly above cutoff


```{r}
pgxFreqplot(freq, chrom = 7)
```
duplications on chromosome 7 are slightly above 20%

```{r}
pgxFreqplot(freq, chrom = 6)
```
deletions on 6q

```{r}
pgxFreqplot(freq, chrom = 9)
```
Very frequent (>25%) duplications on all of chr9

```{r}
pgxFreqplot(freq, chrom =22)
```
Very frequent (>25%) deletions on 22q.


# Analysis

For the cutoff value of 20%, we have most frequent duplications on: 1q, all of 7 and all of chromosome 9. \
and the most frequent deletions are on: 6q and 22q \

If we compare this to the literature, one [study](https://academic.oup.com/neuro-oncology/advance-article/doi/10.1093/neuonc/noad096/7184179) on the subgroup of ependymoma tumors of the posterior fossa group A (PFA) found CNV gains on 1q and losses on 6q.

Another [study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4712639/) found losses of 9p or the total length of chromosome 9.

One [study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC1866851/) also found losses on 22q.