---
title: "BIO392-cnv-freq"
author: ""
date: ""
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Install package

```{r}
```

## Step2: Search esophageal adenocarcinoma NCIt code

## Step3: Access the CNV frequency data from samples with esophageal adenocarcinoma

```{r}
```

### The retreived data is an object contaning two slots `meta` and `data`.

The `meta` slot looks like this:

```{r}
```

The `data` slot includes two matrices.

```{r}
```

The frequency matrix looks like this

```{r}
```

Dimension of this matrix

```{r}
```

## Step4: Visualize data

### By genome

```{r,fig.width=12,fig.height=6}
```

### By chromosome

```{r,fig.width=12, fig.height=6}
```

## Step5: Analyse the data

According the plot, we can see frequenct gains on chromosome 7p, 8q, 20p,20q 
and frequenct losses on chromosome 4p,4q, 5q, 9p, 17p, 18q, 21q.

There is a literature where the findings are consistent with the majority of mine. Here is the [paper-link](https://www.nature.com/articles/modpathol2008150).

A more detailed use case see this [link](https://htmlpreview.github.io/?https://github.com/progenetix/pgxRpi/blob/main/vignettes/Introduction_2_loadfrequency.html).
