---
title: "R-exercise"
author: "Sara Burckhardt"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Basics Exercise Script
Goal: Practice loading data and creating simple visualizations

## 1. Load the Data
Load the `datasets` package (it's built into R, so no need to install it)

```{r}
library(datasets)
```


## 2. Explore the Data
The `iris` dataset is a built-in R dataset with measurements of different flower species

### Display the dataset
 
```{r}
iris
```

### What is the data type of this dataset?    
```{r}
class(iris)
```
  
### Display the first 6 rows of the dataset  
HINT: Use the `head()` function
```{r}
head(iris)
```
  
### Get a summary of the dataset (mean, min, max, etc.)  
```{r}
summary(iris)
```
  
### Get the structure of the dataset (data types, number of rows/columns)  
HINT: Use the `str()` function
```{r}
str(iris)
```

  
## 3. Assign Variables

### Assign petal width and petal length to variables `x` and `y`  
```{r}
x <- c(iris$Petal.Width)
y <- c(iris$Petal.Length)
```


## 4. Visualization

### 4a. Create a scatter plot to display the relationship between petal width and petal length  

```{r}
plot(x,y)
```
    
### 4b. Create a box plot to display Sepal Length for different species  

```{r}
boxplot(Sepal.Length ~ Species, data = iris)
```

    
Try adding your own comments to explain each step!
Feel free to experiment with other plots or modify colors and labels to make the plot more informative!


