---
title: "R Markdown Day 4"
author: "Lea Scherr"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: readable

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Basics Exercise Script

## 1. Load the data
First I loaded the `datasets` package using the `library()` function.

```{r}
library(datasets)
```

## 2. Explore the data
We are using the dataset `iris` in this exercise.

What is the data type of this dataset?
```{r class-of-dataset}
class(iris)
```

Display the first 6 rows of the dataset, to familiarise myself with it.
```{r header}
head(iris)
```

Get a summary of the dataset (mean, min, max, etc.)
```{r summary-table}
summary(iris)
```

Get the structure of the dataset (data types, number of rows/columns)
```{r structure}
str(iris)
```

## 3. Assign the variables
Assign petal width and petal length to variables `x` and `y`.

```{r assign-variables}
x <- iris$Petal.Width
y <- iris$Petal.Length
```

## 4. Visualisation
### Part a) 

Create a scatter plot to display the relationship between petal width and petal length.

```{r scatterplot, fig.width = 4, fig.height = 4}
plot(x, y, col = "darkgreen", 
     xlab = "Petal Width", 
     ylab = "Petal Length")
```

### Part b)

Create a box plot to display Sepal Length for different species.

The function `boxplot` is built like this:  `boxplot(Value ~ Group, data = data)`

```{r boxplot, fig.width = 4, fig.height = 4}
boxplot(Sepal.Length ~ Species, data = iris, 
        xlab = "Species", 
        ylab = "Sepal Length")
```








