---
title: "2021-20-1.BIO392-CNV-freq.mhartmann"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# get the settings ready

```{r}
```{r}
rm(list = ls())
if (!require(devtools)){
    install.packages('devtools')
}
devtools::install_github('progenetix/pgxRpi')
library(pgxRpi)
```
# which cancer? Maybe Thyroid Gland Medullary Carcinoma (Code C3879)

```{r}
freq <- pgxLoader(type='frequency', output='pgxseg',filters='NCIT:C3879',
                  codematches=T)
```


#look at the data :-)
````{r}
freq$meta
freq$data$`NCIT:C3879`
```

#ok I think...

#visualise the data:
#by genome
```{r}
pgxFreqplot(freq)
```

#by chromosome
```{r}
pgxFreqplot(freq, chrom = 21)
```


#There is a loss in 1p, 3p, 3q, 4, 4q, 9p, 9q, 13p and a gain in 1q, 2p, 2q, 5q (19p, 19q)
#[Literture](https://pubmed.ncbi.nlm.nih.gov/33435319/) finds gains in chromosomes 15q, 10q, 14q and 22q and loss in 4q. I do not find this exactly.
#Differetn [Paper](https://academic.oup.com/jcem/article/88/4/1866/2845562?login=true), 4 years later, finds deletion in 1p, 3q, 4, 9q, 13p and 22q and amplifications in 19, #which, I noticed, I find to.
#That is quite that what I find.
#This [study](https://academic.oup.com/jcem/article/99/6/E1104/2537354?login=true) found losses in 1q, 2q and 3q and gians in 2p, 7q, 11q, which I too find.
#So in general, these findings are accepted in the latest literature
