---
title: "CNV in Breast Cancer"
author: "Assignment by Basil Burri"
date: "2025-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **Background and aim of this analysis**

Malignant tumor of breast, also referred to as Malignant Breast Neoplasm, is the
most common global malignancy and the leading cause of cancer deaths in women
[(Katsura et al., 2022)](https://pubmed.ncbi.nlm.nih.gov/35243878/). Copy 
Number Variants (CNVs) contribute significantly to breast cancer susceptibility 
in humans [(Walker et al., 2015](https://pmc.ncbi.nlm.nih.gov/articles/PMC4996380/).
Therefore, this analysis wants to investigate the CNV frequency across the all
chromosomes from breast cancer patients available under the NCIT reference 
"NCIT:C9335" in [Progenetix](https://progenetix.org/subset/?id=NCIT:C9335&datasetIds=progenetix).


### **Preliminary requirements**

#### Install packages

For the analysis, the `pgxRpi` and `GenomicRanges` libraries from 
the `BiocManager` package are required.

BiocManager will ask, whether it should update all your packages.
**Do not update them all. Enter _none_ into the console and press enter.**

```{r eval=TRUE, message=FALSE, warning=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install(version = "3.20", update = FALSE, ask = FALSE)
BiocManager::install(c("pgxRpi", "GenomicRanges"))
```

#### Load required libraries
```{r eval=TRUE, message=FALSE, warning=FALSE}
library(pgxRpi)
library(GenomicRanges)
```


### **Access to CNV data from Progenetix**

#### CNV frequency data from samples with Malignant Breast Neoplasm

Access CNV frequency data from samples with Malignant Breast Neoplasm by using 
the NCIT identifier. `pgxLoader` will access the CNV frequency data of the identifier 
from the [Progenetix](https://progenetix.org/subset/?id=NCIT:C9335&datasetIds=progenetix)
database and assign it to the object.

```{r results=TRUE}
freq <- pgxLoader(type = "cnv_frequency", filter = "NCIT:C9335")
class(freq)
```

#### CNV frequency data from created object

```{r results=FALSE}
freq[["NCIT:C9335"]]
```

#### Show meta data

Use the `mcols` function to show the meta data.

```{r }
mcols(freq)
```


### **Analysis of CNV data**

#### CNV frequency for the whole genome

Identify chromosomes which have duplications and/or deletions based on the 25% 
CNV threshold and could therefore be potential candidates for breast cancer associated CNVs.

**Potential candidates**

| CNV Type           | Chromosomal Arms           |
|--------------------|----------------------------|
| Duplication (Gain) | 1, 8, 11, 16, 17, 20       |
| Deletion (Loss)    | 8, 11, 13, 16, 17, 22      |



```{r fig.width=22, fig.height=6}
pgxFreqplot(freq)
```

### CNV frequency for selected chromosomes

Assess the CNV location on the chromosomes which are potential candidates 
for breast cancer associated CNVs.

```{r fig.width=15, fig.height=6}
pgxFreqplot(freq, chrom=c(1, 8, 11, 13, 16, 17, 20, 22), layout = c(2, 4))

```

### Results


**Findings**

Assessed arms on the chromosomes (p: small, q: long) where the duplications 
and/or deletions are located based on th 25% CNV threshold.

| CNV Type           | Chromosomal Arms                 |
|--------------------|----------------------------------|
| Duplication (Gain) | 1q, 8q, 11q, 16p, 17q, 20q       |
| Deletion (Loss)    | 8p, 11q, 13q, 16q, 17p, 22q      |



**Comparison to literature**

A [Genome-wide Copy Number Alteration study](https://pmc.ncbi.nlm.nih.gov/articles/PMC3224104/) 
lists the most frequent (frequency >40% in patients) Copy Number Alterations (CNAs) for various breast cancer subtypes.
The table below annotates all found CNVs of this analysis based on the results from the study.

| Chromosomal Arm | CNV Type           | Subtypes with High Frequency (>40%)                      |
|-----------------|--------------------|----------------------------------------------------------|
| 1q              | Duplication        | ER-positive, ER-negative, HER2-positive, Triple Negative |
| 8q              | Duplication        | ER-positive, ER-negative, HER2-positive, Triple Negative |
| 11q             | Duplication        | None                                                     |
| 16p             | Duplication        | None                                                     |
| 17q             | Duplication        | HER2-positive                                            |
| 20q             | Duplication        | None                                                     |
| 8p              | Deletion           | ER-positive, ER-negative, HER2-positive, Triple Negative |
| 11q             | Deletion           | ER-positive, ER-negative, HER2-positive, Triple Negative |
| 13q             | Deletion           | Triple Negative                                          |
| 16q             | Deletion           | ER-positive, ER-negative, HER2-positive, Triple Negative |
| 17p             | Deletion           | HER2-positive                                            |
| 22q             | Deletion           | ER-negative, HER2-positive                               | 



**Discussion**

This analysis found three chromosomal locations of copy number variation 
that are not associated to  breast cancer subtypes based on the 
[study](https://pmc.ncbi.nlm.nih.gov/articles/PMC3224104/). This mismatch 
could be explained by the strict 25% threshold that was used in this analysis. When we 
investigate the degree of copy number variation for the duplications on 
11q, 16p and 2q, we can see that the 25% threshold for duplication was only slightly exceeded.
This mismatch suggests that the strict threshold does not mirror the clinical phenotype of CNVs.
By applying the 25% threshold, this analysis was able to select chromosomes with CNVs that
are potentially associated to breast cancer. This result suggest that we can use the 25% threshold
for preliminary selection of potential candidates which can get evaluated afterwards by referring
to studies that assessed the frequency of CNVs in patients with certain disease.  


