---
title: "BIO 392 - CNV.Freq"
author: "Eric Thiele"
date: "2023-09-27"
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Tumor Information

Uncontrolled proliferation and spread of abnormal cells within the
breast tissue is a characteristic of invasive breast cancer. In more
advanced stages, these malignant cells may travell through the
bloodstream or lymphatic system to distant regions of the body in
addition to having the capacity to invade neighbouring tissues. Breast
cancer most often occurs in women between 50 to 74, but can also occur
in men.

# Invasive Breast Carcinoma

The NCIt code is C9245

# Access the CNV frequency data

-   load library

```{r}
library(pgxRpi)
```

# Access data

```{r}
freq <- pgxLoader(type='frequency', output='pgxfreq', filters='NCIT:C9245')
```

-   The metadata information is

```{r}
freq$meta
```

-   The data looks like this

```{r}
freq$data$`NCIT:C9245`[c(1:3),]

```

# Visualisation

```{r}
pgxFreqplot(freq)

```

\ 
The given bars show the CN increase of a chromosome, whereas the blue
ones indicate CN loss relative to the whole reference genome.\

```{r}
pgxFreqplot(freq, chrom=1)
```

  There is a clear gain of CN in the q-part of the chromosome 1. The
gain of the CN's in Chromosome 1 often correlates with breast
carcinoma.\

```{r}
pgxFreqplot(freq, chrom=8)
```

 In the 8th chromosome, you can see on the one end deletions as well as
a gain of CN's on the longer q-arm.\

```{r}
pgxFreqplot(freq, chrom=16)
```

 Chromosome 16 is a clear case for deletions of CN's.

# Data Analysis

Chromosome gains and deletions for a 25% threshold are shown here.

According to the plot, there are several gains of CN's on the\
chromosom:\
1q, 8q, 16p, 17q, 20q\

And several deletions on chromosom:\
1p, 4pq, 6q, 8p, 9pq, 11q, 13q, 15q, 16q, 17p, 18q, 22q, Xpq

In the Study: [Genomic alterations during the in situ to invasive ductal
breast carcinoma transition shaped by the immune
system][<https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8026652/>]\
we compare ceratin deletions and gains of the CN's.\

From the Study you get:\
Amplification: 1q, 8q, 17q, 20q\
Deletions: 8p, 11q, 16q, 17p\

Comparing from the study with the library\

The overlaps for duplication are:\
1q, 8q, 17q, 20q\

As well for the deletionsare:\
8p, 11q, 16q, 17p\

# conclusion

It can be said that the chromosomes 1q, 8q, 17q, 20q
-gains and 8p, 11q, 16q, 17p - deletions correlate with breast cancer
with a very high probability, as they are found in the Protogenix
library on the one hand, as well as being mentioned in an independent
study.
