---
title: "bio392-cnv-freq"
author: "Luca Tobler"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # dont use cashe
```

# Tumor Information: non-small cell carcinoma
Non-small cell lung cancer is a type of epithelial lung cancer. It is the most common type of lung cancer, accounting for 85% of all cases. The most common types of NSCLC are squamous cell carcinoma, large cell carcinoma, and adenocarcinoma, but there are several other types that occur less frequently. Its NCit code is C2926. 

## Load the data
```{r}
library(pgxRpi)
freq <- pgxLoader(type="frequency",output="pgxfreq",filters="NCIT:C2926")
```

## The metadata information is:
```{r}
freq$meta
```


# the data looks like this: 
```{r}
freq$data$"NCIT:C2926"[c(1:5),]
```


# Visualisation
```{r}
pgxFreqplot(freq)
```

```{r}
pgxFreqplot(freq,chrom=c(8,18)) 
```

# Analysis

According to the plot, frequent copy number gains occur on 1q, 5p, chr7, 8q, while frequent loss of copy numbers occur here: 3p, 6q, 8p, 9p, 13p, 15q, 17p, chr18 and 19p. 

This [study](https://onlinelibrary.wiley.com/doi/10.1002/gcc.22460) found most gains to occur on 2p, 12p, and parts of 22q. These do unfortunately not overlap with the results from NCit data. The study found looses on 6q, 15q, 19p, and 22q, which does overlap with the NCit data (overlaps on 6p, 15q and 19p). Issues may be that the study focuses on further classifying non-small cell lung carcinoma into lung adenocarcinoma (LUAD) and lung squamous cell carcinoma (LUSC) and therefore focuses its samples on these two subtypes, while the data we used from NCit includes other subtypes as well. 