---
title: "bio392-cnv-freq"
author: "Solange Jost"
date: "27 9 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Invasive breast carcinoma

Invasive breast carcinoma is a type of breast cancer that has spread into surrounding breast tissue. The most common subtypes include invasive ductal carcinoma, invasive lobular carcinoma and inflammatory breast cancer.
The NCIt code for invasive breast carcinoma is C9245.

## Access the CNV frequency data

### load library

```{r}
library(pgxRpi)
```

### access data

```{r}
freq <- pgxLoader(type='frequency', output='pgxfreq', filters="NCIT:C9245")
```

The metadata information is

```{r}
freq$meta
```

There are 3'796 samples labelled with Invasive Breast Carcinoma.


The data looks like this:

```{r}
freq$data$'NCIT:C9245'[c(1:3),]
```

## Visualization

```{r}
pgxFreqplot(freq)

```

We look at some interesting chromosomes in more detail to decide if the regions meet our set criteria of being above 25%:

```{r}
pgxFreqplot(freq, chrom=9)
pgxFreqplot(freq, chrom=16)
pgxFreqplot(freq, chrom=20)
```

We count the whole chromosome 9 as a loss/deletion, chr16 and chr20 don't quite meet our criteria (one of the criteria being that a whole chromosome arm must be above the threshold of 25%).

## Analysis

Looking at our graph, we have frequent chromosomal gain on 1q and 8q as well as frequent chromosomal loss on 4, 9, 13q, 18 and X.

Looking up a [paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8026652/) on invasive breast carcinoma, copy number aberrations mentioned are 1q, 8q, 17q and 20q amplification as well as 8p, 11q, 16q and 17p loss. 

Comparing these results to our data, every single one is clearly visible on our graph and would have been included in our interpretation as well if our criteria were chosen differently.
Note that none of our stated loss regions (4, 9, 13q, 18 and X) are mentioned in the paper.

